### 2016-04-12 1kG/workflow.Rmd
thousand genomes init

---
title: "Local PCA on 1000 genomes"
date: "`r date()`"
---

```{r setup, include=FALSE}
library(devtools)
load_all("../package")
# Chromosome map lengths, from:
.chrlens <- c( 262.00830, 244.03647, 209.97332, 197.48005, 189.58412, 174.70526, 172.35610, 156.52810, 146.28861, 160.40336, 143.88007, 156.40816, 119.34065, 104.83476, 110.93523, 118.68929, 119.61966, 105.74726, 92.57134, 83.62057, 54.79314, 55.59191 )
.chrstarts <- c(0, cumsum(.chrlens) )
# and in bp
.chrbps <- as.numeric( c( 247135059, 242663303, 199318155, 191167888, 180625439, 170747902, 158798338, 146264218, 140185941, 135284541, 134449982, 132276874, 114092980, 106356482, 100210760, 88684276, 78605474, 76115554, 63785051, 62376958, 46902240, 49522492 ) )
.chrbpstarts <- c(0,cumsum(.chrbps))
```

# The data

We have VCFs from *phase 3* of the 1000 genomes, in separate VCF files by chromosome,
from 2504 individuals.
Of these, 24 are relatives; we'll remove these.
```{r vcfs}
vcfs <- list.files("phase3",".*vcf.gz$", full.names=TRUE)
full.sample.info <- read.csv("20130606_sample_info.csv",header=TRUE,stringsAsFactors=FALSE)
vcf.header <- scan( pipe( paste0( "zcat ", vcfs[1], " | head -n 1000 | grep '^#[^#]'")), what='char' )
vcf.indivs <- vcf.header[-(1:9)]
all( vcf.indivs %in% full.sample.info$Sample )
sample.info <- subset(full.sample.info, Sample %in% vcf.indivs)
rellies <- unlist(subset(sample.info, ! ( Relationship %in% c("", "unrels", "unrel", "not father")) )[,7:14])
rellies <- unlist( strsplit( rellies[nchar(rellies)>0], "[ ,]" ) )
included.rellies <- intersect( rellies, vcf.indivs )
sample.info$usethese <- ! ( sample.info$Sample %in% included.rellies )
table(subset(sample.info,usethese)$Population)
```

Now, we would like to find the eigendecomposition
of each genomic window.
We'll start with chromosome 22.
```{r chr22, cache=TRUE}
vcf.chr22 <- grep("chr22",vcfs,value=TRUE)
# first do a scan to find the scale we're dealing with
coords <- scan( pipe( paste0( "zcat ", vcf.chr22, " | grep -v '^#' | cut -f 2 " ) ) )
length(coords)
summary( diff(coords) )
# percent of spacings less than 100bp
mean(diff(coords)<100)
length(coords)
```
On chromsome 22, which is 1.7% of the genome, there are 1.1 million SNPs,
median spacing between SNPs is 20bp, and 95% of spacings are below 100bp.