### 2016-05-05 medicago/local_pca_on_medicago.Rmd
rewrote eigen_windows to take functions

# Local PCA on *Medicago truncatula* Hapmap

```{r setup}
devtools::load_all("../package")
```

# Data

Filtered SNP calls from Mt4.0 were downloaded from [medicagohapmap.org](http://medicagohapmap.org/) on 5/4/2016:
```
mkdir -p data; cd data
wget http://www.medicagohapmap.org/downloads/Mt40/Mt4.0_HapMap_README.pdf
for CHROM in 1 2 3 4 5 6 7 8 u;
do
    echo "Getting $CHROM"
    wget http://www.medicagohapmap.org/downloads/Mt40/snps_by_chr/chr${CHROM}-filtered-set-2014Apr15.bcf
    wget http://www.medicagohapmap.org/downloads/Mt40/snps_by_chr/chr${CHROM}-filtered-set-2014Apr15.csi
done
```
For some reason these .csi index files don't work; so we need to remake them:
```
for CHROM in 2 3 4 5 6 7 8 u;
do
    bcftools index -f chr${CHROM}-filtered-set-2014Apr15.bcf
done
```

## Samples

The file of sample info was obtained from the HTML table at [http://www.medicagohapmap.org/hapmap/germplasm](http://www.medicagohapmap.org/hapmap/germplasm).
To check this contains information about all samples,
first get the list of samples in the .bcf files:
```
bcftools query -l chr1-filtered-set-2014Apr15.bcf > samples_in_bcf.txt
```
then compare to the table of information:
```{r sample_info}
samples <- read.table("data/sample_info.tsv", sep="\t", header=TRUE)
bcf.samples <- scan("data/samples_in_bcf.txt", what='char')
setdiff( bcf.samples , samples$ID )
```
For some reason, the bcf files refer to `HM020-I` while in the the sample info file there is `HM020`.
We'll assume they are the same (but keep an eye on that one!).
Here's where they're all from:
```{r show_sample_info}
bcf.samples <- gsub("-I","",bcf.samples)
samples <- droplevels(subset(samples, ID %in% bcf.samples))
sort( table(samples$Country.of.Origin) )
```

## Recoding

Recall that VCF files have the following columns:

> CHROM  POS     ID      REF     ALT     QUAL    FILTER  INFO    FORMAT

We use [bcftools](https://github.com/samtools/bcftools) to extract information, for instance:
```
# list of sites
bcftools query -f '%CHROM\t%POS\n' -r chr1:1-400 chr1-filtered-set-2014Apr15.bcf
# genotypes
bcftools query -f '%CHROM\t%POS[\t%GT]\n' -r chr1:1-400 chr1-filtered-set-2014Apr15.bcf
# genotypes of first four individuals, numerically
bcftools query -f '%CHROM\t%POS[\t%GT]\n' -r chr1:1-400 chr1-filtered-set-2014Apr15.bcf -s 'HM001,HM002,HM003,HM004' | sed -e 's_0/0_0_g' -e 's_\(0/1\)\|\(1/0\)_1_g' -e 's_1/1_2_g'
```

To get windows out of the VCF file and recode them, we'll just pipe a call to `bcftools`.
Here's a dummy version that will pull out windows of 100 SNPs for only the first four individuals.
```{r win_fn}
get.indivs <- c("HM001","HM002","HM003","HM004")
bcf.file <- "data/chr1-filtered-set-2014Apr15.bcf"
# first read in the info about sites
bcf.con <- pipe(paste("bcftools query -f '%POS\\n' -r chr1:1-100000",bcf.file),open="r")
bcf.sites <- list( chr1=scan(bcf.con) )
close(bcf.con)
# now use bcftools to pull out sites we want
win.fn <- function (n,win.snps=100) {
    sites <- list( chr1=bcf.sites[[1]][seq(1+(n-1)*win.snps,length.out=win.snps)] )
    regions <- region_string(sites)
    query_genotypes( bcf.file, regions, get.indivs )
}
win.fn(20,10)  # get the 20th window of 10 SNPs
```

Here's an alternate version that pulls out windows based on basepairs.
Note that this version assumes we have only one chromosome, starting at zero.
```{r win_fn_bp}
win.fn.bp <- function (n,win.bp=200) {
    win.start <- 1+(n-1)*win.bp
    win.end <- win.start + win.bp-1
    regions <- paste0( "chr1", ":", win.start, ",", win.end )
    query_genotypes( bcf.file, regions, get.indivs )
}
win.fn.bp(20,50)
```

## Computing eigenvectors of covariance matrices

Now we are set up to use `lostruct::eigen_windows()` to compute and write out the eigenstructure for windows along the genome.

Here's how to do that for the first 4 windows of length 100 SNPs, with the simplified version: just four individuals.
```{r simple_snp_pcas}
eigen_windows( 
        data=bcf.file,
        win.fn=function(data,n,...){ win.fn(n,...) },
        do.windows=1:4,
        k=2,
        win.snps=100
    )
```
Here's the same thing on windows of length 1000bps:
```{r simple_bp_pcas}
eigen_windows( 
        data=bcf.file,
        win.fn=function(data,n,...){ win.fn.bp(n,...) },
        do.windows=1:4,
        k=2,
        win.bp=1000
    )
```

## Doing this for everything

Now let's see how it goes to do this for an entire chromosome.
On chromosome 1 there are about 5 million SNPs, so we'll do it in windows of 5,000.
First, windows by SNPs.
```{r lpca_chr1}
sites <- vcf_positions(bcf.file)
win.fn.snp <- vcf_windower(bcf.file, size=5e3, type="snp", sites=sites, samples=get.indivs) 
```