### 2016-05-11 medicago/medicago_data_setup.Rmd
run_on_medicago.R runs, saves regions and labeled MDS coords

# Local PCA on *Medicago truncatula* Hapmap

```{r setup}
devtools::load_all("../package")
library(Matrix)
library(colorspace)
options(width=100)
fig.dim <- 4
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
```

```

# Data

Filtered SNP calls from Mt4.0 were downloaded from [medicagohapmap.org](http://medicagohapmap.org/) on 5/4/2016:
```
mkdir -p data; cd data
wget http://www.medicagohapmap.org/downloads/Mt40/Mt4.0_HapMap_README.pdf
for CHROM in 1 2 3 4 5 6 7 8 u;
do
    echo "Getting $CHROM"
    wget http://www.medicagohapmap.org/downloads/Mt40/snps_by_chr/chr${CHROM}-filtered-set-2014Apr15.bcf
    wget http://www.medicagohapmap.org/downloads/Mt40/snps_by_chr/chr${CHROM}-filtered-set-2014Apr15.csi
done
```
For some reason these .csi index files don't work; so we need to remake them:
```
for CHROM in 2 3 4 5 6 7 8 u;
do
    bcftools index -f chr${CHROM}-filtered-set-2014Apr15.bcf
done
```

## Samples

The file of sample info was obtained from the HTML table at [http://www.medicagohapmap.org/hapmap/germplasm](http://www.medicagohapmap.org/hapmap/germplasm).
To check this contains information about all samples,
first get the list of samples in the .bcf files:
```
bcftools query -l chr1-filtered-set-2014Apr15.bcf > samples_in_bcf.txt
```
then compare to the table of information:
```{r sample_info}
samples <- read.table("data/sample_info.tsv", sep="\t", header=TRUE,stringsAsFactors=FALSE)
bcf.samples <- scan("data/samples_in_bcf.txt", what='char')
setdiff( bcf.samples , samples$ID )
```
For some reason, the bcf files refer to `HM020-I` while in the the sample info file there is `HM020`.
We'll assume they are the same (but keep an eye on that one!).
Here's where they're all from:
```{r show_sample_info}
wierd.names <- grep("-I",bcf.samples,value=TRUE)
samples$ID[ match(gsub("-I","",wierd.names),samples$ID) ] <- wierd.names
samples <- droplevels(subset(samples, ID %in% bcf.samples))
sort( table(samples$Country.of.Origin) )
```

## Statistics by chromosome

We should know what the density of SNPs looks like, for choosing window sizes.
Chromosome lengths are from the VCF files.
```{r snp_density, cache=TRUE}
chrom.lens <- c( chr1=52991155, chr2=45729672, chr3=55515152, chr4=56582383, chr5=43630510, chr6=35275713, chr7=49172423, chr8=45569985, chl_Mt=124033 )
chrom.stats <- sapply( paste0("chr",1:8), function (chrom) {
            bcf.file <- file.path("data",sprintf("%s-filtered-set-2014Apr15.bcf",chrom))
            sites <- vcf_positions(bcf.file)
            spacings <- diff(sites[[chrom]])
            out <- c( nsnps = length(sites[[chrom]]),
                    nbp = chrom.lens[chrom],
                    spacing.mean = mean(spacings),
                    spacing = quantile(spacings,.05),
                    spacing = quantile(spacings,.25),
                    spacing = quantile(spacings,.50),
                    spacing = quantile(spacings,.75),
                    spacing = quantile(spacings,.95),
                    spacing.max = max(spacings,.95) )
            return(out)
        } )
floor(chrom.stats)
```

So, thats a SNP every 10bp.
Since we have `r length(bcf.samples)` samples,
each local PC will take `r 1+length(bcf.samples)` doubles to record;
a double is 8 bytes;
so in, say 1Gb we can fit one local PC for `r floor(1e9 / (8*(1+length(bcf.samples))))` windows.
With a total of `r sum(chrom.stats['nsnps',])` SNPs,
that'd be `r floor(sum(chrom.stats['nsnps',])/floor(1e9 / (8*(1+length(bcf.samples)))))` SNPs per window.
A bigger concern is the distance matrix we compute between windows
(we don't have to save this to disk, though).
With this many windows, that would be `r floor(1e9 / (8*(1+length(bcf.samples))))^2` floats,
or `r floor(1e9 / (8*(1+length(bcf.samples))))^2 / 8e9` Gb.
That's still possible, but barely.

## Recoding

Recall that VCF files have the following columns:

> CHROM  POS     ID      REF     ALT     QUAL    FILTER  INFO    FORMAT

We use [bcftools](https://github.com/samtools/bcftools) to extract information, for instance:
```
# list of sites
bcftools query -f '%CHROM\t%POS\n' -r chr1:1-400 chr1-filtered-set-2014Apr15.bcf
# genotypes
bcftools query -f '%CHROM\t%POS[\t%GT]\n' -r chr1:1-400 chr1-filtered-set-2014Apr15.bcf
# genotypes of first four individuals, numerically
bcftools query -f '%CHROM\t%POS[\t%GT]\n' -r chr1:1-400 chr1-filtered-set-2014Apr15.bcf -s 'HM001,HM002,HM003,HM004' | sed -e 's_0/0_0_g' -e 's_\(0/1\)\|\(1/0\)_1_g' -e 's_1/1_2_g'
```

To get windows out of the VCF file and recode them, we'll just pipe a call to `bcftools`.
Here's a dummy version that will pull out windows of 100 SNPs for only the first four individuals.
```{r win_fn}
get.indivs <- c("HM001","HM002","HM003","HM004")
bcf.file <- "data/chr1-filtered-set-2014Apr15.bcf"
# first read in the info about sites
bcf.con <- pipe(paste("bcftools query -f '%POS\\n' -r chr1:1-100000",bcf.file),open="r")
bcf.sites <- list( chr1=scan(bcf.con) )
close(bcf.con)
# now use bcftools to pull out sites we want
win.fn <- function (n,win.snps=100) {
    sites <- list( chr1=bcf.sites[[1]][seq(1+(n-1)*win.snps,length.out=win.snps)] )
    regions <- enclosing_region(sites)
    vcf_query( bcf.file, regions, get.indivs )
}
win.fn(20,10)  # get the 20th window of 10 SNPs
```

Here's an alternate version that pulls out windows based on basepairs.
Note that this version assumes we have only one chromosome, starting at zero.
```{r win_fn_bp}
win.fn.bp <- function (n,win.bp=200) {
    win.start <- 1+(n-1)*win.bp
    win.end <- win.start + win.bp-1
    regions <- data.frame( chrom="chr1", start=win.start, end=win.end )
    vcf_query( bcf.file, regions, get.indivs )
}
win.fn.bp(20,50)
```

## Computing eigenvectors of covariance matrices

Now we are set up to use `lostruct::eigen_windows()` to compute and write out the eigenstructure for windows along the genome.

Here's how to do that for the first 4 windows of length 100 SNPs, with the simplified version: just four individuals.
```{r simple_snp_pcas}
eigen_windows( 
        data=win.fn,
        do.windows=1:4,
        k=2,
        win.snps=100
    )
```
Here's the same thing on windows of length 1000bps:
```{r simple_bp_pcas}
eigen_windows( 
        data=win.fn.bp,
        do.windows=1:4,
        k=2,
        win.bp=1000
    )
```

## Doing this for everything

Now let's see how it goes to do this for an entire chromosome.
On chromosome 1 there are about 5 million SNPs, so we'll do it in windows of 5,000.
First, windows by SNPs.
```{r lpca_chr1_snp, cache=TRUE}
sites <- vcf_positions(bcf.file)
do.indivs <- sample( samples$ID, 50 )
win.fn.snp <- vcf_windower(bcf.file, size=5e3, type="snp", sites=sites, samples=do.indivs) 
system.time( snp.pca <- eigen_windows(win.fn.snp,k=2) )
```
It takes about 2 minutes to do chromosome 1, for ten individuals, on phoebe.
Next, we compute the distance between windows with `pc_dist()` and do MDS visualization on the result.
This is just with ten samples, so we don't expect a real result, but let's check it all works:
```{r mds_chr1_snp, cache=TRUE, depends="lpca_chr1_snp"}
system.time( chr1.pcdist <- pc_dist( snp.pca ) )
# there may be windows with missing data
na.inds <- is.na( snp.pca$values[1,] )
chr1.mds <- cmdscale( chr1.pcdist[!na.inds,!na.inds], eig=TRUE, k=4 )
mds.coords <- chr1.mds$points[ ifelse( na.inds, NA, cumsum(!na.inds) ), ]
colnames(mds.coords) <- paste("MDS coordinate", 1:ncol(mds.coords))
```

## Visualization

Here's the results:
```{r show_results_snp}
# the distance matrix itself
image( Matrix( chr1.pcdist ) )
# the MDS results
win.regions <- region(win.fn.snp)()
win.mids <- (win.regions$start+win.regions$end)/2
chrom.cols <- rainbow_hcl(128, c=90, end=.9*360)[as.numeric(cut(win.mids,128))]
pairs( mds.coords, pch=20, col=adjustcolor(chrom.cols,0.75) )
```

Now let's extract the extreme windows in the MDS plot:
```{r get_corners}
mincirc <- lostruct:::enclosing_circle( mds.coords[,1:2] )
mds.corners <- corners( mds.coords[,1:2], prop=.05 )
corner.cols <- c("red","blue","purple")
ccols <- rep("black",nrow(mds.coords))
for (k in 1:ncol(mds.corners)) {
    ccols[ mds.corners[,k] ] <- corner.cols[k]
}
plot( mds.coords[,1:2], pch=20, col=adjustcolor(ccols,0.75), 
        xlab="MDS coordinate 1", ylab="MDS coordinate 2", 
        xlim=mincirc$ctr[1]+c(-1,1)*mincirc$rad,
        ylim=mincirc$ctr[2]+c(-1,1)*mincirc$rad )
plot_circle( mincirc, col='red' )
points( mincirc$three, col='red', cex=2 )
points( mds.coords[mincirc$index,], col='red', cex=1.5 )
```

Here is where these lie along the chromosomes:
```{r show_corners}
for (k in 1:ncol(mds.coords)) {
    plot( win.mids/1e6, mds.coords[,k], pch=20, 
            xlab="Position (Mb)", ylab=paste("MDS coordinate",k),
            col=adjustcolor(ccols,0.75) )
}
```

And, here are the PCA plots corresponding to these corners:
```{r pc_corners, cache=TRUE, depends="mds_chr1_snp"}
corner.winfn <- function (n) { win.fn.snp( mds.corners[,n] ) }
corner.pcs <- eigen_windows( data=corner.winfn, 
                    k=2, do.windows=1:ncol(mds.corners) )
```

```{r plot_pc_corners}
countries <- samples$Country.of.Origin[match(do.indivs,samples$ID)]
country.names <- unique(countries)
country.cols <- rainbow_hcl( length(country.names), c=90 )
country.pch <- rep(1:6,length.out=length(country.names))
for (k in 1:ncol(corner.pcs$vectors)) {
    xy <- matrix( corner.pcs$vectors[,k], nrow=nrow(corner.pcs$vectors)/nrow(corner.pcs$values) )
    plot(xy, xlab="PC 1", ylab="PC 2", main=paste("corner",k), 
            col=country.cols[match(countries,country.names)], 
            pch=country.pch[match(countries,country.names)]  )
}
legend('topright', col=country.cols, pch=country.pch, legend=country.names)
```