### 2016-05-30 medicago/summarize_run.Rmd
.

---
title: "Local PCA results"
date: "`r date()`"
---

```{r setup, include=FALSE}
library(lostruct)
library(colorspace)
library(jsonlite)
library(RColorBrewer)
fig.dim <- 4
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
```

Render this, for instance, like:
```
templater::render_template("summarize_run.Rmd",output="lostruct/results_type_snp_size_10000_jobid_324902/run_summary.html",change.rootdir=TRUE)
```

```{r data_setup, include=FALSE}
opt <- fromJSON("config.json")
if (is.null(opt$weights)) { opt$weights <- 1 }

# original data files
chroms <- paste0("chr",1:8)
bcf.files <- file.path( "..", "..", "data", paste0(chroms,"-filtered-set-2014Apr15.bcf") )
names(bcf.files) <- chroms

sample.ids <- vcf_samples(bcf.files[1])
samp.file <- file.path( "..", "..", "data", "sample_info.tsv" )
samps <- read.table(samp.file,sep="\t",header=TRUE)
samps <- droplevels( samps[match(gsub("-I$","",sample.ids),samps$ID),] )

# data files, precomputed
pca.files <- list.files(".",".*.pca.csv")
mds.file <- "mds_coords.csv"
regions.files <- list.files(".",".*.regions.csv")

# read in mds
mds.coords <- read.csv(mds.file,header=TRUE)
mds.cols <- (1:ncol(mds.coords))[-(1:2)]

# position information
regions <- do.call( rbind, lapply( regions.files, read.csv, header=TRUE ) )
# figure out where to plot things at
chroms <- unique(regions$chrom)
chrom.starts <- tapply( regions$start, regions$chrom, min, na.rm=TRUE )
chrom.ends <- tapply( regions$end, regions$chrom, max, na.rm=TRUE )
chrom.spacing <- floor(.05*mean(chrom.ends))
chrom.offsets <- c(0,cumsum(chrom.spacing+chrom.ends))
names(chrom.offsets) <- c(chroms,"end")
chrom.dividers <- c(0,chrom.offsets[-1])-chrom.spacing/2
chrom.mids <- chrom.dividers[-1] - diff(chrom.dividers)/2

# this is where to plot windows at when plotting with all chromosomes
regions$pos <- chrom.offsets[regions$chrom]+(regions$start+regions$end)/2

chrom.cols <- rainbow_hcl(length(chroms), c=90, end=.9*360)[as.numeric(regions$chrom)]

#' Set up to plot all chromosomes together
#' will plot the vector of values 'y' along all the chromosomes.
chrom.plot <- function (y,ylab='',main='',chrom.labels=TRUE,...) {
    plot(0, type='n', xlim=range(chrom.offsets/1e6), ylim=range(y,finite=TRUE), 
         xlab='', xaxt='n', ylab=ylab, main=main)
    for (k in 1:floor(length(chroms)/2)) {
        rect( xleft=chrom.dividers[2*k-1]/1e6, xright=chrom.dividers[2*k]/1e6, 
             ybottom=par("usr")[3], ytop=par("usr")[4], 
             border=NA, col=adjustcolor("grey",0.25) )
    }
    abline( v=chrom.dividers/1e6, lty=3, col=adjustcolor("grey",0.5), lwd=2 )
    if (chrom.labels) axis( 1, at=chrom.mids/1e6, labels=chroms, las=0, tick=FALSE )
    points( regions$pos/1e6, y, ...)
}
```

This run had these parameters:

- *window type* : `r opt$type`
- *window size* : `r opt$size`
- *number of pcs* : `r opt$npc`
- *number of MDS coordinates* : `r opt$nmds`

Here are the number of windows per chromsome,
and the computed MDS coordinates, colored by chromosome:
```{r mds_pairplot, fig.height=2*fig.dim}
table(regions$chrom)
pairs( mds.coords[,mds.cols], pch=20, col=adjustcolor(chrom.cols,0.75) )
```

Here are the extreme windows in the MDS plot:
```{r get_corners}
mds.corners <- corners( mds.coords[,mds.cols[1:2],drop=FALSE], prop=.05 )
corner.coords <- t( apply( mds.corners, 2, function (k) { colMeans( mds.coords[k,mds.cols,drop=FALSE], na.rm=TRUE ) } ) )
# set up colors and pchs for corners
corner.cols <- brewer.pal(3,"Dark2")
corner.pch <- c(15,17,19)
ccols <- rep("black",nrow(mds.coords))
cpch <- rep(20,nrow(mds.coords))
for (k in 1:ncol(mds.corners)) {
    ccols[ mds.corners[,k] ] <- corner.cols[k]
    cpch[ mds.corners[,k] ] <- corner.pch[k]
}
```
```{r plot_corners, fig.width=3*fig.dim, fig.height=fig.dim*ceiling(length(mds.cols)/2)}
# plot corners and MDS along the chromosome
layout(matrix(c(rep(1,length(mds.cols)),1+seq_along(mds.cols)),ncol=2),
        widths=c(1,2), heights=c(rep(1,length(mds.cols)-1),1.25))
plot( mds.coords[,mds.cols[1:2]], pch=cpch, 
     col=adjustcolor(ccols,0.75),  asp=1,
     xlab="MDS coordinate 1", ylab="MDS coordinate 2" )
text( corner.coords, labels=1:3 )
opar <- par(mar=c(par("mar"),1.1)[c(5,2,5,4)])
for (k in mds.cols) {
    lastone <- (k==mds.cols[length(mds.cols)])
    if (lastone) {
        par(mar=c(opar$mar,0.2)[c(1,2,5,4)])
    }
    chrom.plot( mds.coords[,k], pch=20, 
            xlab=if (lastone) { "Position (Mb)"} else { "" }, # main=paste("MDS coordinate",match(k,mds.cols)),
            chrom.labels=lastone,
            ylab=colnames(mds.coords)[k],
            col=adjustcolor(ccols,0.75) )
}
par(opar)
```

Now, we'll look at PCA plots from the extracted corners.
(this is done without a ton of memory by accumulating the covariance matrix in `running_cov`):
```{r corner_pca, cache=TRUE}
corner.npc <- 4
corner.regions <- lapply( 1:ncol(mds.corners), function (k) {
            regions[ mds.corners[,k],]
    } )
corner.covmats <- lapply( 1:ncol(mds.corners), function (k) {
            reg <- regions[ mds.corners[,k], ]
            qfun <- multi_vcf_query_fn( chrom.list=chroms, file=bcf.files, regions=reg )
            running_cov(qfun,1:nrow(reg))
        } )
corner.pca <- lapply( corner.covmats, function (covmat) {
            cov_pca(covmat=covmat,k=corner.npc,w=opt$weights)
    } )
```

Here is the color scheme:
```{r pop_colors, fig.height=2*fig.dim}
pop.names <- levels(samps$Country.of.Origin)
pop.cols <- rainbow_hcl(nlevels(samps$Country.of.Origin))
pop.pch <- seq_len(nlevels(samps$Country.of.Origin))
plot( rep(1,length(pop.names)), seq_along(pop.names), pch=pop.pch, col=pop.cols, xlim=c(0,length(pop.names)),
       xlab='', ylab='', xaxt='n', yaxt='n' )
text( rep(1,length(pop.names)), seq_along(pop.names), labels=pop.names, pos=4 )
```

Here are all pairwise plots of the first `r corner.npc` PCs for each of the three corners:
```{r plot_corner_pca, fig.width=3*fig.dim}
layout(t(1:3))
for (i in 1:(corner.npc-1)) {
    for (j in (i+1):corner.npc) {
        for (k in 1:ncol(mds.corners)) {
            vectors <- matrix( corner.pca[[k]][-(1:(1+corner.npc))], ncol=corner.npc )[,c(i,j)]
            colnames(vectors) <- paste("PC", c(i,j))
            plot(vectors, pch=pop.pch[samps$Country.of.Origin], 
                    col=pop.cols[samps$Country.of.Origin], cex.axis=1.5 )
            if (i==1 && j==2) {
                mtext(paste("corner",k),side=3,cex=2,col=corner.cols[k])
            }
        }
    }
}
```